Leaf mass per area, not total leaf area, drives differences in above-ground biomass distribution among woody plant functional types

### 2016-04-27 manuscript.Rnw
Final changes - version as submitted

% Notes:
% - For submission, convert to Word docx using Latex2rtf. 
% - Do not include line numbers here, add them in Word. Also do not embed figures. 
% - Word count can be automatically added on first page; follow instructions further below.

@@

% -- Word count
% After compiling the ms (and no cleanup, so .tex is present), run:
% shell("texcount manuscript.tex > wordcount.txt")
% Then uncomment the following:
%' <<echo=FALSE, eval=FALSE>>=
%' nrw <- get_wordcount("wordcount.txt")
%' @
%' Total number of words: \Sexpr{nrw$Total} \\
%' Abstract: \Sexpr{nrw$Abstract} \\
%' Introduction: \Sexpr{nrw$Introduction} \\
%' Methods and Materials: \Sexpr{nrw$Methods} \\
%' Results: \Sexpr{nrw$Results} \\
%' Discussion: \Sexpr{nrw$Discussion} \\
%' Number of figures: 6 (all in colour) \\
%' Number of tables: 2 \\
%' Supporting information : 6 Figures (all in colour).

@@

A final aim was to test for shifts in biomass distribution with climate. Little is known about global-scale patterns in biomass distribution in relation to climate, but one might expect biomass distribution to shift with precipitation or mean annual temperature, as small-scale comparisons have shown such responses \citep{berninger1994,callaway1994,delucia2000,poyatos2007}. \citet{reich2014} has also shown that after controlling  for the large size-related variation in biomass distribution, the ratio of leaf to total plant mass (averaged across whole stands, either including or excluding below-ground components) increases with mean annual temperature in both gymnosperms and angiosperms. Thus, we aimed to test for climate-related variation in $\mfms$ and $\afms$, and their component terms with PFTs.

@@

Similar to \citet{poorter2015}, we found that none of the studied relationships were linear on a log-log scale. Therefore, we did not use standard allometric equations to study relationships between size and biomass variables. Instead, we used generalised additive models (GAM), a semi-parametric modelling approach that makes no assumption about the shape of the relationships, to visualise the relationships between biomass and plant size variables, and to estimate variables such as $\mfms$ and $\afas$ at a common plant height. In all fitted GAMs, we used a cubic regression spline. For the smoothed term in the model (plant height), we used up to 3 or 4 degrees of freedom, which resulted in biologically realistic smoothed relationships. Within the GAM, we used a penalised regression smoother \citep{wood2006} to allow the final degree of smoothness to be estimated from the data. In all fitted GAMs, we used species-dataset combination as a random effect. To homogenise variance, all variables (except MAP and MAT) were log-transformed prior to analysis.

@@

>>>>

% Notes:
% - For submission, convert to Word docx using Latex2rtf:
%   - set clean=FALSE in call to latex_build() in remake.yml. Then run, 'latex2rt manuscript.tex' (in shell)
% - Do not include line numbers here, add them in Word. Also do not embed figures. 
% - Word count can be automatically added on first page; follow instructions further below.

@@

% -- Word count
% After compiling the ms (and no cleanup, so .tex is present), run:
% shell("texcount manuscript.tex > wordcount.txt")
% Then uncomment the following:
<<echo=FALSE, eval=TRUE>>=
nrw <- get_wordcount("wordcount.txt")
@
Total number of words: \Sexpr{nrw$Total} \\
Abstract: \Sexpr{nrw$Abstract} \\
Introduction: \Sexpr{nrw$Introduction} \\
Methods and Materials: \Sexpr{nrw$Methods} \\
Results: \Sexpr{nrw$Results} \\
Discussion: \Sexpr{nrw$Discussion} \\
Number of figures: 6 (all in colour) \\
Number of tables: 2 \\
Supporting information : 6 Figures (all in colour).

@@

A final aim was to test for shifts in biomass distribution with climate. Little is known about global-scale patterns in biomass distribution in relation to climate, but one might expect biomass distribution to shift with precipitation or mean annual temperature, as small-scale comparisons have shown such responses \citep{berninger1995,callaway1994,delucia2000,poyatos2007}. \citet{reich2014} has also shown that after controlling  for the large size-related variation in biomass distribution, the ratio of leaf to total plant mass (averaged across whole stands, either including or excluding below-ground components) increases with mean annual temperature in both gymnosperms and angiosperms. Thus, we aimed to test for climate-related variation in $\mfms$ and $\afms$, and their component terms with PFTs.

@@

Similar to \citet{poorter2015}, we found that none of the studied relationships were linear on a log-log scale. Therefore, we did not use standard allometric equations to study relationships between size and biomass variables. Instead, we used generalised additive models (GAM), a semi-parametric modelling approach that makes no assumption about the shape of the relationships, to visualise the relationships between biomass and plant size variables, and to estimate variables such as $\mfms$ and $\afms$ at a common plant height. In all fitted GAMs, we used a cubic regression spline. For the smoothed term in the model (plant height), we used up to 3 or 4 degrees of freedom, which resulted in biologically realistic smoothed relationships. Within the GAM, we used a penalised regression smoother \citep{wood2006} to allow the final degree of smoothness to be estimated from the data. In all fitted GAMs, we used species-dataset combination as a random effect. To homogenise variance, all variables (except MAP and MAT) were log-transformed prior to analysis.

@@

The result that $\afas$ did not correlate with climate seems to contradict reports, from studies on one or few species, that the ratio of sapwood area to leaf area (the `Huber value`) varies across climate gradients as quantified by summer evaporative demand \citep{delucia2000,poyatos2007}, rainfall and temperature \citep{callaway1994}, and temperature \citep{berninger1995}. Possible reasons for the discrepancy include 1) we studied $\as$ (stem basal area) including the sapwood and heartwood, 2) small-scale climate effects may have been present in our dataset but average out on large climate gradients, and 3) climate effects may be highly species-specific; the majority of reported correlations between the Huber value and climate were for species in the genus \emph{Pinus}.
